---
title: "IPL"
author: "Nikhil Pillai"
date: "22 September 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(ggplot2)
library(corrplot)
library(fmsb)

```

##Reading the data
```{r}
ipl = read.csv("C:/Users/Administrator/Desktop/EDA/IPL/IPL.csv")
```


### Dimension and structure of data 
```{r}
dim(ipl)
str(ipl)
```


### Percentage of missing values in each columns  
```{r}
na.percent = colSums(is.na(ipl))/nrow(ipl)*100
na.percent
```

### There are no null values present in the data set 


### Percentage of data availibility in each column
```{r}
data_avalibility_percent = colSums(!is.na(ipl))/nrow(ipl)*100
data_avalibility_percent
```

### The code shows 100 % data whereas there are missing values in the columns: Player_dismissed, dismissal_kind and fielder.  


### Univariant analysis (Dimensional columns)
```{r}
col_name = "batting_team"
frequency = ipl %>% group_by(Group=get(col_name)) %>% summarise(Total = n(), percent = n()/nrow(ipl)*100) 
kable(frequency)
ggplot(frequency, aes(x=Group,y=percent)) + geom_bar(stat='identity') + coord_flip() + theme_bw()
```


### Univariant analysis (numerical columns)
```{r}
hist(ipl$total_runs)
boxplot(ipl$total_runs)
summary(ipl$total_runs)
```


### Total Runs vs attributes 
```{r}
attribute_name = "batting_team"
bivariant_analysis = ipl %>% group_by(Group = get(attribute_name)) %>% summarise(Total_Runs=sum(total_runs)) 
View(bivariant_analysis)
ggplot(bivariant_analysis, aes(x=Group, y=Total_Runs)) + geom_bar(stat='identity') + coord_flip() + theme_bw()
```


### Top batsman against runs scored
```{r}

Batsman_runs = ipl %>% group_by(batsman) %>% summarise(Total_Runs=sum(total_runs)) %>% arrange(-Total_Runs)
Top_batsman = head(Batsman_runs,10)
Top_batsman
ggplot(Top_batsman, aes(x=batsman, y=Total_Runs)) + geom_bar(stat='identity') + coord_flip() + theme_bw()
```

### Bowlers given most number of runs 
```{r}
bowler_runs = ipl %>% group_by(bowler) %>% summarise(Total_Runs=sum(total_runs)) %>% arrange(-Total_Runs)
Top_bowlers = head(bowler_runs,10)
Top_bowlers
ggplot(Top_bowlers, aes(x=bowler, y=Total_Runs)) + geom_bar(stat='identity') + coord_flip() + theme_bw()
```


### PCA
```{r}
ipl_pca = ipl[,c("wide_runs","bye_runs","legbye_runs","noball_runs","penalty_runs","batsman_runs","extra_runs","total_runs")]

eigen_vectors = eigen(cor(ipl_pca))$vectors
eigen_values = eigen(cor(ipl_pca))$values 

dim(ipl_pca)
dim(eigen_vectors)

ipl_pcs = scale(ipl_pca) %*% eigen_vectors
```


### Cross check eqivalence of pcs variance and eigen values
```{r}
print(var(ipl_pcs[,1]))
print(eigen_values[1])
```


### Cross check correlation matrix
```{r}
corr_ipl_pcs = cor(ipl_pcs)
corrplot(corr_ipl_pcs, method = "number")
```


### Identifying the number of columns to ignore 
```{r}
eigen_values/length(eigen_values)*100 

vars_cum = cumsum(eigen_values/ length(eigen_values)*100)
vars_cum

cor(ipl_pcs[,2], ipl_pcs[,3])
```

#### The correlation between the pc's will be nearly zero as all pc's are orthogonal to each other 
#### Each pc's has an impact on the output but the same impact is there in other pc's
#### We can infer that the last three columns compute to 100% accuracy 


##Getting dates
```{r, include=FALSE}
ipl = ipl %>% mutate(date = as.Date(date, "%d-%m-%Y"))
ipl = ipl %>% mutate(Day = format(date, "%d"), Month = format(date, "%B"), Year = format(date, "%Y"), Week = weekdays(date))
```

```{r, include=FALSE}

ipl = ipl %>% mutate(Wide_runs_count = if_else(wide_runs == 0, 0, 1))
ipl = ipl %>% mutate(Bye_runs_count = if_else(bye_runs == 0, 0, 1))
ipl = ipl %>% mutate(Legbye_runs_count = if_else(legbye_runs == 0, 0, 1))
ipl = ipl %>% mutate(Noball_runs_count = if_else(noball_runs == 0, 0, 1))
ipl = ipl %>% mutate(Penalty_runs_count = if_else(penalty_runs == 0, 0, 1)) 
ipl = ipl %>% mutate(Batsman_runs_count = if_else(batsman_runs == 0, 0, 1))
ipl = ipl %>% mutate(Total_runs_count = if_else(total_runs == 0, 0, 1))
ipl = ipl %>% mutate(Sixes = if_else(batsman_runs == 6, 1, 0))
ipl = ipl %>% mutate(Fours = if_else(batsman_runs == 4, 1, 0))
ipl = ipl %>% mutate(Wickets = if_else(dismissal_kind == '', 0, 1))

```


##Yearwise Purple Cap Winners.
```{r}
ipl = ipl %>%
  mutate(Wickets = if_else(dismissal_kind == '' , 0, 1))

purple_cap = ipl %>%
  group_by(Year, bowler) %>%
  summarise(No_of_wickets = sum(Wickets)) %>%
  arrange(Year) %>%
  filter(No_of_wickets == max(No_of_wickets))

ggplot(purple_cap, aes(x = Year, y = No_of_wickets, fill = bowler)) + geom_bar(stat = 'identity') + theme_bw() + labs(title = 'Purple Cap Winners')
```

##Yearwise Orange Cap Winners.
```{r}
orange_cap = ipl %>%
  group_by(Year,batsman) %>%
  summarise(Runs_scored = sum(total_runs)) %>%
  arrange(Year) %>%
  filter(Runs_scored == max(Runs_scored))

ggplot(orange_cap, aes(x = Year, y = Runs_scored, fill = batsman)) + geom_bar(stat = 'identity') + theme_bw() + labs(title = 'Orange Cap Winners')

lmn = ipl %>%
  group_by(Year,batting_team, batsman) %>%
  summarise(Runs_scored = sum(total_runs)) %>%
  arrange(Year, -Runs_scored) %>%
  filter(Runs_scored == max(Runs_scored))

```

##Average runs per year
```{r}
No_of_teams = ipl %>%
  group_by(Year) %>%
  summarise(Teams = length(unique(batting_team)))

ipl2 = ipl %>%
  group_by(Year) %>%
  mutate(Teams = length(unique(batting_team)))

Avg_runs_team = ipl2 %>%
  group_by(Year) %>%
  summarise(Runs = sum(total_runs), Avg_runs = sum(total_runs)/max(Teams))

kable(Avg_runs_team)
```

##Run distribution of top ten run scorers by year
```{r}
player_runs = ipl %>%
  group_by(batsman) %>%
  summarise(Runs = sum(total_runs)) %>%
  arrange(-Runs)

player_runs = head(player_runs, 10)
Top_ten_players = ipl[ipl$batsman %in% player_runs$batsman, ]

ggplot(Top_ten_players, aes(x = Year, y = total_runs)) + geom_bar(fill = 'navy', stat = 'identity') + facet_wrap(~batsman) + theme(axis.text.x = element_text(angle = 90)) + labs(title = 'Run distribution of top ten run scorers by year')

```

##Average runs per team
```{r}
abc = ipl %>% 
  group_by(match_id, batting_team, Year) %>% 
  summarise(team_score = sum(total_runs)) %>% 
  arrange(match_id)

winning_team = abc %>% 
  group_by(match_id, Year) %>% 
  filter(team_score == max(team_score))

wins = winning_team %>%
  group_by(batting_team) %>%
  summarise(No_of_wins = n()) %>%
  arrange(-No_of_wins)


four_teams = head(wins, 4)
Top_four_teams = ipl[ipl$batting_team %in% four_teams$batting_team | ipl$bowling_team %in% four_teams$batting_team, ]

Total_runs = ipl %>%
  group_by(batting_team) %>%
  summarise(Runs = sum(total_runs)/length(unique(match_id))) %>%
  arrange(-Runs)

```

```{r}
ggplot(Total_runs, aes(x = batting_team, y = Runs, group = 1)) + geom_bar(stat = 'identity', fill = 'red') + theme(axis.text.x = element_text(angle = 90))
```



##Total runs by over
```{r}
runs = ipl %>% 
  group_by(over) %>% 
  summarise(total_runs_over = sum(batsman_runs)) %>%
  arrange(-total_runs_over)
kable(runs)
```



##Teamwise batting and bowling average
```{r}
Teams_best1 = ipl %>% group_by(batting_team) %>% summarise(Runs = sum(total_runs),Sixes = sum(Sixes),Fours = sum(Fours),Strike_rate = sum(total_runs)/n()*100)

Teams_best2 = ipl %>% group_by(bowling_team) %>% summarise(Wick = sum(Wickets),Bowl_avg = sum(total_runs)/sum(Wickets))

Teams_best = cbind(Teams_best1,Teams_best2)
Teams_best = Teams_best %>%
  mutate(Batting_avg = Runs/Wick)

Teams_best5 = Teams_best %>%
  select(-Runs, -Wick) %>%
  arrange(-Strike_rate)


ggplot(Teams_best5, aes(x = batting_team, group = 1)) + geom_line(aes(y = Bowl_avg)) + geom_line(aes(y = Batting_avg), col = 'red') + theme(axis.text.x = element_text(angle = 90))
```



##Ball wise avg runs

```{r}
balls = ipl %>% group_by(ball) %>% summarise(avg_runs = mean(batsman_runs), count = n()) %>% arrange(-count)
kable(balls)
```


#total number of extra runs given in that ball

```{r}
extra = ipl %>% group_by(ball) %>% summarise(total_extra_runs = sum(extra_runs))
kable(extra)
```

####According to this, the 4th ball is the best as the least number of extra runs have been given to the batting team.


##Total number of runs given in that ball due to no ball

```{r}
noball = ipl %>% group_by(ball) %>% summarise(total_runs_noballs = sum(noball_runs), sum(Noball_runs_count))
kable(noball)
```

## Total number of times a player has been responsible for a run out.

```{r}
player_run_out = ipl %>% filter(dismissal_kind == 'run out') %>% group_by(fielder) %>% summarise(count = n()) %>% arrange(-count)
kable(player_run_out[1:10, ])
```



##Players who have retired due to getting hurt.

```{r}
player_hurt = ipl %>% filter(dismissal_kind == 'retired hurt') %>% group_by(batsman, batting_team) %>% summarise(count = n()) %>% arrange(-count)
kable(player_hurt)
```

####There have been only 8 retire hurts in the history of ipl, 6 of whom have been from mumbai indians.



```{r}
pqr = ipl %>%
  group_by(batting_team, Year) %>%
  summarise(Sum = sum(total_runs))

ggplot(pqr, aes(x = Year, y = Sum)) + geom_bar(fill = 'orange', stat = 'identity') + facet_wrap(~batting_team) + theme(axis.text.x = element_text(angle = 90)) 
```











